Automated Verification Applied to Stand-alone Solar Photovoltaic Systems: Optimal Sizing and Project Validation

### 2019-07-03 Tex/Old text/abbreviations.tex
thesis updated text

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Renewable \& Sustainable Energy Reviews}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

%%%%%packages to makecell, algorithm, not inclueded with Elsevier
\usepackage{makecell}
\usepackage{algorithm,algorithmic}
%\algsetup{linenosize=\small}
\graphicspath{ {./images/} }
\usepackage{amsmath}
\usepackage{xcolor}

\begin{document}

\begin{frontmatter}

%\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
\title{Automated Formal Verification of Stand-alone Solar Photovoltaic Systems}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
%\author{Alessandro Trindade\fnref{myfootnote}, Lucas Cordeiro}
%\address{Radarweg 29, Amsterdam}

%% or include affiliations in footnotes:
\author[mymainaddress]{Alessandro Trindade\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{alessandrotrindade@ufam.edu.br}
%\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Lucas Cordeiro}

\address[mymainaddress]{Federal University of Amazonas, Av. Rodrigo Octávio, 6200, Coroado I, 69077-000 Manaus-AM-Brazil}
\address[mysecondaryaddress]{University of Manchester, School of Computer Science, Kilburn Building, Manchester M13 9PL}

\begin{abstract}
With declining costs and increasing performance, the deployment of renewable energy systems is growing faster. In 2017, for the first time, the number of people without access to electricity dipped below 1 billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design space. Automated verification using model checking has proven to be an effective technique to any kind of program verification. This paper marks the first application of software model checking to formally verify the design of a stand-alone solar photovoltaic system including solar panel, charge controller, battery, inverter, and electric load. Case studies, from real photovoltaic systems deployed in five different sites, ranging from 700W to 1,200W, were used to evaluate this proposed approach and to compare that with specialized simulation tools. Different verification tools were evaluated as well, in order to compare performance and soundness. Data from practical applications show the effectiveness of our approach, where specific conditions that lead to failures in a photovoltaic solar system are only detailed by the automated verification method.
\end{abstract}

\begin{keyword}
Formal verification \sep model checking \sep photovoltaic power systems \sep power system modeling \sep solar power generation
%\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
%\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\section*{List of Abbreviations, Units and Nomenclature}
\noindent $a$ Ideality or quality factor \\
AC Alternating current \\
$C_{10}$ 10h of rated battery capacity \\
$C_{bank}$ Total capacity of the battery bank \\
CBMC C Bounded Model Checker \\
$ cos \varphi $ Power factor of the inverter \\
CPA Configurable Program Analysis \\
CPAchecker Configurable Program Analysis Checker \\
DC Direct current \\
$DOD$ Depth of discharge ($DOC=1-SOC$) \\
$E_{consumption}$ Energy consumption \\
$E_{corrected}$ Energy corrected \\
$E_{p}$ Energy produced by one panel \\
ESBMC Efficient SMT-based Bounded Model Checker \\
$G$ Solar irradiance \\
$ G_{ref} $ Solar irradiance at STC ($1000 W/m^{2}$) \\
$I$ Current \\
$I_{0}$ Reverse saturation current \\
$ I_{AC} $ Current that the inverter supply to the house \\
$I_{c,min}$ Charge controller minimum current \\
$ I_{DC} $ Current required by the inverter from the DC source \\
$ I_{in} $ Input current \\
$I_{load}$ Current that feed the load \\
$ I_{out} $ Output current \\
$I_{ph}$ Photocurrent \\
$ I_{ph,ref} $ Photocurrent at STC \\
$I_{pv}$ Current from the PV system \\
$ I_{sc} $ Short-circuit current \\
$ I_{sc,amb} $ Ambient short-circuit current \\
$ I_{sc,ref} $ Short-circuit current at STC \\
$ k_{B} $ Boltzmann constant ($ 1.3806503\times10^{-23}J/K $); \\
$MAX_{AC,ref}$ Peak power that the inverter can support \\
$N$ Number of series-connected cells \\
$N_{B}total$ Sized number of batteries \\
$N_{BP}$ Sized number of parallel batteries \\
$N_{BP}min$ Total minimum number of parallel batteries \\
$N_{BS}$ Sized number of series batteries \\
$N_{BS}min$ Total minimum number of series batteries \\
NOCT Nominal Operating Cell Temperature \\
$N_{PS}$ Sized number of panels in series \\
$N_{PSmin}$ Total number of panels in series \\
$N_{PP}$ Sized number of panels in parallel \\
$N_{PPmin}$ Total umber of panels in parallel \\
$ N_{TP} $ Sized number of panels \\
$N_{TPmin}$ Total minimum number of needed solar panels \\
PV Photovoltaic \\
P Power \\
$P_{in}$ Input power \\
$P_{out}$ Output power \\
$P_{surge}$ Maximum peak power demanded by house for few seconds \\
$q$ Absolute value of the electron's charge ($ -1.60217646\times10^{-19}C $) \\
SAT Satisfiability or Boolean satisfiability problem \\
SAT When returns a failure from the verification tool \\
SMT Satisfiability modulo theories \\
$ SOC $ Battery state of charge \\
STC Standard Test Condition (reference state) \\
$T$ Temperature \\
$ T_{air} $ Ambient temperature \\
$ T_{ref}$ Temperature at STC ($=298.15 K$ or $25^{o}C$) \\
UNSAT Condition when verification tool succeeds (could not reach an error) \\
V Voltage \\
$ V_{AC}  $ Voltage that the inverter supply to the house \\
$V_{c}$ Battery charge voltage  \\
$V_{d}$ Battery discharge voltage \\
$ V_{DC} $ Input voltage to the inverter delivered by the DC source \\
$ V_{in} $ Input voltage \\
$ V_{oc} $ Open-circuit voltage \\
$ V_{oc,ref} $ Reference open-circuit voltage at STC \\
$ V_{out} $ Output voltage \\
$ V_{system} $ DC voltage of the bus \\
$V_{T}$ Thermal voltage ($ V_{T}=k_{B}T/q $) \\
$\Delta T$ Temperature variation ($\Delta T=T-T_{ref} $) \\
$ \mu_{I} $ Short-circuit current temperature coefficient ($A/K$) \\
$ \mu_{V} $ Open-circuit voltage temperature coefficient ($ V/K $) \\
$\eta_{b}$ Battery efficiency \\
$\eta_{c}$ Charge controller efficiency \\
$\eta_{i}$ Inverter efficiency \\
$\eta_{p}$ PV panel efficiency \\

%\linenumbers

\section{Introduction}
Lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}. Progress has been made worldwide; in particular, the number of people without electricity access fell below 1 billion threshold for the first time in 2017~\cite{IEAweo2018}. The share of people without access to electricity from Africa is 58\%, while 19\% of the share comes from developing Asia, and 31\% from Latin America~\cite{IEAweo2018}. Numbers from Brazil show the aim to electrify 270 isolated places and 2.7 million of people until 2023~\cite{EPE2018}. Furthermore, there is a close relationship between the lack of energy and the low HDI (Human Development Index) of those localities~\cite{Coelho}.

In order to provide universal electricity for all, decentralized systems led by solar photovoltaic (PV) in off-grid and mini-grid systems will be the lowest-cost solution for three-quarters of the additional connections needed; and grid extension will be the standard especially in urban areas~\cite{IEAweo2018}.

%The definition of energy access usually includes both electricity access and access to modern fuels for cooking and heating, in order to replace traditional biomass. 
%Increased access to energy allows economic growth and poverty alleviation \cite{Karekesi}. 
%
%Among the options of renewable sources, there are hydro, wind, and solar photovoltaic (PV). 
Only a niche market a few years ago, solar photovoltaic (PV) systems are now becoming a mainstream electricity provider, with an increase of approximately 50\% from 2016 to 2017 in terms of new installations of PV~\cite{EPIA}. However, this growth is not driven by isolated systems.

In order to simulate or evaluate a PV system there are a myriad of specialized tools available in the market, such as RETScreen, HOMER, PVWatts, SAM, and Hybrid2 \cite{Pradhan,Swarnkar,NRELDobos,NRELBlair,Mills}; and even general purpose simulation tools such as PSpice, Saber, or MATLAB/Simulink package \cite{Gow1999,Benatiallah2017}.
%
%However, to the best of our knowledge, this work can be the first work to perform automated verification of a stand-alone solar PV project solution.  
%
%According to \cite{SEIA} and \cite{Chauhan}, solar is the most abundant source of renewable energy on earth. In SPV system, solar radiations are captured from the sun and turned into electricity using SPV cells made of silicon and other materials. SPV...
%
%In order to address different aspects of a SPV project, there is a myriad of public domain and commercial software available. According to \cite{Brooks}, the capabilities of these tools range from simple solar resource and energy production estimative, to site survey and system design tools, to complex financial analysis software (with optimization). Some tools also provide support to rebate programs applications and tax incentives (specific to each country or region), while other programs and worksheets focus on the technical aspects of system sizing and design.  
%
%Manufactures and integrators have yet their proprietary software to perform various system sizing, with the drawback of include just their own products among the possibilities of choice, what restrict the solution. 
%
%PV systems can be classified into grid-connected and stand-alone systems. At this proposal, only off-grid or stand-alone systems will be considered. The utilization of solar energy in off-grid mode has the potential to meet the energy need for remote rural areas of developing countries.  
However, those tools are based on running experiments in simulation models. Simulation has the advantage of being cheap (if compared to test in real systems) and can be employed before the system design is concluded but it has the drawback of an incomplete coverage since the verification of all possible combinations and potential failures of a system is unfeasible~\cite{ClarkeHV18}.

Formal methods based on model checking offer a great potential to obtain a more effective and faster verification in the design process~\cite{ClarkeHV18}.  
%with behavior that is, in principle, well determined. 
Any kind of system can be specified as computer programs using mathematical logic, which constitutes the intended (correct) behavior; then, one can try to give a formal proof or otherwise establish that the program meets its specification. %User or project requirements can be added during the creation of the formal model to be verified. %, as defined by \cite{Trindade}. 
%
%This area of research is referred as formal methods. % as defined in \cite{Clarkeetal}.
% Their aim is to establish system correctness with mathematical rigor. 
%In the recent decades, research in formal methods has led to the development of very promising verification techniques.
% that facilitate the early detection of bugs in order to ensure the correctness of the system. 
%
%Model-based verification techniques are based on models that describe the possible system behavior in a mathematically precise and unambiguous manner. 
%Thus, such problems as incompleteness, ambiguities, and inconsistencies, which normally are discovered only in later stages of the design, can be detected in advance, as described by \cite{Trindade}. 
%Model checking algorithms can then verify the system model by systematically exploring all its states to check whether the requirements are met by the given system.
%
%Formal methods, alternatively, provide facilities for the modeling of each component of any complex system~\cite{Hall1990}. 
%
In this study, a mathematical model of each component of a stand-alone PV system, as panel solar, charge controller, batteries, inverter, and electrical load is created. The behavior of each system component can be analyzed and observed with the support of those formal models, as a joint operation of the components, which in this case represents the operation of the solar PV system itself. The project requirements, as battery autonomy and power demand, besides weather conditions, as solar irradiance and ambient temperature, are translated as part of the computer program and automatically verified during the formal process. The model checking tool reports in which conditions a system does not meet the user requirements. A key benefit to this approach is that it helps in the detection of flaws in the design phase of system development, thereby considerably improving system reliability~\cite{Akram2018}. 
%Those programs are viewed as mathematical objects, and the use of 
%Bringing this technique to solar PV systems, i
%Instead of finding a bug or a software inconsistence, t
%Here it will be verified a stand-alone PV system, with all their components modeled mathematically,   
%Among the non-deterministic variables considerated, is it worth to mention environmental issues (like solar irradiance, and temperature), and the variation of the electrical load (power consumption) or the battery' state of charge, as shown in Fig.\ref{fig:flowchartgeneral}.
The implementation of the proposed tool is carried out by means of an algorithm written in language C, that is executed by three state-of-the-art model checkers to formally verifying PV designs, in order to evaluate performance and correctness: the C Bounded Model Checker (CBMC)~\cite{Kroening}, the Efficient SMT-based Bounded Model Checker (ESBMC)~\cite{esbmc2018}, and the Configurable Program Analysis Checker (CPAchecker).
%three the efficient SMT-based bounded model checker (ESBMC)~\cite{esbmc2018}. %, which allows one to incrementally verify a PV system as an imperative program using a fragment of decidable first-order theories~\cite{DBLP:books/daglib/0019162}.

In prior studies, the evaluation of PV systems w.r.t. user requirements were performed by simulation tools using MATLAB/Simulink~\cite{Benatiallah2017, Natsheh2012}, or HOMER Pro \cite{Lamnadi2017}. Related to formal methods, studies were carried out toward the modeling of power smart grids~\cite{Akram2018} and solar panels~\cite{Driouich2017}; however, those studies do not perform automated verification. 
%
In addition, recent research performed a formal study of PV farms, where the focus has been the interaction with the grid~\cite{Abate2017}; or to model a PV system and to verify the maximum power point of solar panels with the use of Jmodelica tool~\cite{Driouich2018}. Both studies restrict to PV panels, and do not include batteries, inverters, and charge controllers.

%The experimental results show that the proposed framework is more efficient \textcolor{red}{than what?} in different case studies to find not proper functioning of PV systems than simulation with off-the-shelf tool. 
%This is the first study to apply a formal verification technique to formally check the design of a stand-alone solar PV system. In summary, t
This paper makes three original contributions. % Firstly, we describe a modular modeling of each component of a PV system by means of mathematical models that can be encoded into fragments of first-order theories supported by software model checkers. Secondly
Firstly, we propose an algorithm written in language C that implements the automated verification method which formally checks the sizing and the operation of a given stand-alone PV system. Secondly, we evaluate the verification method by comparing three state-of-the-art model checkers in five real case studies. Thirdly, experimental results show that this proposed approach can find subtle design errors in stand-alone PV systems, which are not easily detected by other approaches based on simulation.

%\textcolor{red}{This introduction is missing the description of the vision, approach and results. In particular, try to match each paragraph to: problem, vision, approach, results and contributions.} 
%
%The framework provided at this paper outlines the requirements, the process and the mathematical modeling to perform automated formal verification of a stand-alone PV system, using the satisfiability modulo theories (SMT)-based verification method.

\textit{Outline}. Section~\ref{sec:SolarPhotovoltaicSystem} gives the background about solar PV systems, design and validation of PV systems, and the mathematical modeling. Section~\ref{sec:AutomatedVerification} presents the automated verification technique. Section~\ref{sec:sizingcheck} shows how to perform the sizing of a stand-alone PV system. The methodology is presented in section~\ref{sec:Methodology}. Section~\ref{sec:results} is devoted to the results. Section~\ref{sec:Conclusions} presents the conclusion and describes future work.

%--------------------------------------------------------
\section{Solar Photovoltaic System }
\label{sec:SolarPhotovoltaicSystem}
%--------------------------------------------------------

%According to \cite{Roy}, a 
%A PV system is designed to give the electric supply to a load. This load can be either Alternate Current (AC) or Direct Current (DC) type. Electric supply can be needed either in daytime or evening time. 
%(in particular cases, in both times). %The most basic PV system can give supply only in daytime.  
%For night hours or rainy days, one needs to hold batteries, where power can be stored and used.% \cite{Gules}.

PV systems are classified into distinct types~\cite{Mohanty}. %: (1) stand-alone systems, where the energy is generated and consumed in the same place and which do not interact with the main grid; (2) grid-connected systems; and (3) solar PV hybrid system. %, where another source(s) of energy can work together with the solar PV. % system to provide the required demand.
%, such as wind, biomass or diesel, 
%As the goal of this work is just the solutions aimed to isolated/ off-grid application, %therefore it is not considered the on-grid or hybrid configurations.
%
%\begin{itemize}
%\item Stand-alone systems, where the energy is generated and consumed in the same place and which does not interact with the main grid. Normally, the electricity consuming/utilizing device is part of the system, i.e., solar home systems, solar street lighting system, solar lanterns, and solar power plants; 
%\item Grid-connected systems, where the solar PV system is connected to the grid. The grid-connected system can be either a grid-tied system, which can only feed power into the grid and such system cannot deliver power locally during blackouts and emergencies since these systems have to be completely disconnected from the grid and have to be shut down as per national and international electrical safety standards. Some grid-connected PV systems, with energy storage, can also provide power locally in an islanding mode; 
%\item Solar PV hybrid system: In a hybrid system, another source(s) of energy, such as wind, biomass or diesel, can work together with the solar PV system to provide the required demand. In such type of system, main objective is to bring more reliability into the overall system at an affordable way by adding one or more energy sources.
%\end{itemize}
Specifically for remote rural areas of developing countries or places where the grid extension is not feasible, 
%stand-alone systems are one the most used solutions, and 
% depending on the type of load, cost, resources availability, and requirements of the load, 
%stand-alone systems 
%can be split into several categories. However, 
the most suitable configuration is the regulated stand-alone system with battery and AC load; this configuration is the focus of this study.
%, which is the target of this paper, includes PV array, batteries, charge controller, inverter, and AC load.
%PV array, battery, MPPT and AC load. Battery is used to store the extra power of PV system. A charge controller is necessary for this type of system because the useful life of the battery is less than that of the PV module. Extra charging and deep discharging can reduce the battery life \cite{Kim}. Because of the AC load, an inverter (DC to AC converter) is required. 
%This solution has an increase of cost because has more equipment. However, the AC availability brings the advantage of a higher number of AC appliances to use at the houses or consumer units.
%The resource MPPT, called of maximum power point tracking, is an electronic control mechanism that maintains the PV operating in a voltage that correspond to the voltage of maximum power.
%, which maximizes the transfer of power and avoiding lost at the PV cells \cite{Pinho}. %That resource is found at modern PV systems and is strongly indicated due to its advantages.

%\subsection{Unregulated stand-alone PV system with DC load }
%Usually this type of system is for low power applications, as defined by \cite{Roy}. The PV system is directly connected to the load without any MPPT controller, as shown at Fig.\ref{fig:standaloneSPV} letter (a). At night hours, the system will not provide any supply because of the absence of the battery.

%\begin{figure}[h]
%\includegraphics[width=0.5\textwidth]{reg_unreg4.png}
%\centering
%\caption{PV stand-alone systems: (a) regulated stand-alone SPV system with DC load; (b) regulated stand-alone SPV system with DC load; (c) regulated stand-alone SPV system with battery DC load; and (d) regulated stand-alone system with battery and AC load  (Source: \cite{Roy}).}
%\label{fig:standaloneSPV}
%\end{figure}

%\subsection{Regulated stand-alone PV system with DC load}
%It is similar to unregulated stand-alone system with DC load, but the main difference between this and the previous one is that this system requires a MPPT technique, as illustrated by Fig.\ref{fig:standaloneSPV} letter (b). Usually system with MPPT should have battery; otherwise, extra power will be waste.

%\subsection{Regulated stand-alone system with battery and DC load}
%Configuration with PV array, battery, MPPT and DC load, as shown in Fig.\ref{fig:standaloneSPV} letter (c). Battery is used to store the extra power of PV system. A charge controller is necessary for this type of system because the useful life of the battery is less than that of the PV module. Extra charging and deep discharging can reduce the battery life \cite{Kim}.

%According to \cite{Pinho}, PV systems that are used to feed loads with low variation on the consumption can be sized to operate without the controller. That is called of self-regulated stand-alone PV system with battery. However, the voltage from the PV panel must be compatible with the batteries voltage. Normally, the bank of batteries are oversized related to the PV panel and to the load. The drawback is the operation of the batteries, normally overloaded or with excessive discharges (that can damage the batteries).

%\subsection{Regulated stand-alone system with battery and AC load}
%This system is similar to the previous one, but here AC load draws the power from PV system and, because of the AC load, an inverter (DC to AC converter) is required, as seen in Fig.\ref{fig:standaloneSPV} letter (d). This solution has an increase of cost because has more equipments. However, the AC availability brings the advantage of a higher number of AC appliances to use at the houses or consumer units. That configuration is the focus of the present work since it is the most common nowadays all over the world, when off-grid or isolated regions is the target.

\subsection{Design and Simulation of Solar PV systems}
%The design and validation of a PV system can be done by hand or with the aid of a software tool. 
%
%In order to address different aspects of the PV system design, there are various software tools available in the literature~\cite{Rajanna,Rawat}.
%public domain and commercial software available for the PV market. 
%According to \cite{Brooks}, t
%%%%%%%%%%The capabilities of tools available in the literature range from simple solar resource and energy production estimation, %to site survey and system design tools,
% to complex financial analysis and project optimization. 
%Some tools also provide support to rebate programs applications and tax incentives (specific to each country or region), while other programs and worksheets focus on the technical aspects of system sizing and design.
%
%Manufacturers and integrators have yet their proprietary software to perform inverter string sizing and various system sizing and design tools, with the drawback to just include their own products among the possibilities of choice. 
Here we evaluated the most popular tools available in the literature and summarized at Table~\ref{table:softwares}: PVWatts, SAM, HOMER, RETScreen, and Hybrid2~\cite{Pradhan,Swarnkar,NRELDobos,NRELBlair,Mills}.

%\subsection{PVWatts Calculator}
%According to \cite{Freeman} and \cite{NRELDobos}, it is a web application developed by the National Renewable Energy Laboratory (NREL), which estimates the electricity production of a grid-connected roof- or ground-mounted photovoltaic system based on a few inputs. According to \cite{NRELDobos}, to use the calculator, it is necessary to provide information about the system's location, basic design parameters, and system economics. PVWatts calculates estimated values for the system's annual and monthly electricity production, and for the electricity monetary value. This tool is suitable for very preliminary studies of a potential location for a photovoltaic system that uses crystalline silicon or thin film photovoltaic modules. The production estimates that PVWatts computation do not account for many factors that are important in the design of a photovoltaic system, therefore it is necessary support of an energy expert. The calculator estimates the monthly and annual electricity production of a photovoltaic system using an hour-by-hour simulation over a period of one year. To represent the system's physical characteristics, PVWatts requires values for six inputs: System DC size; Module type; Array type; System losses; Array tilt angle; Array azimuth angle.
%
%\subsection{SAM}
%SAM or System Advisor Model is a software from the U.S. Department of Energy and National Renewable Energy Laboratory. According to \cite{NRELBlair} and \cite{Cameron2008}, SAM is intended to help users to determine whether the model meets their project constraints/specifications, and (also) to provide information for readers who do not plan to use the model, but want to learn about its capabilities. SAM is a performance and financial model designed to facilitate decision making for people involved in the renewable energy industry: Project managers and engineers; Financial and policy analysts; Technology developers; and Researchers. SAM makes performance predictions and cost of energy estimates for grid-connected power projects based on installation and operating costs and system design parameters that you specify as inputs to the model. Projects can be either on the customer side of the utility meter, where they buy and sell electricity at retail rates, or on the utility side of the meter, where they sell electricity at a price negotiated through a power purchase agreement. SAM is an electric power generation model and assumes that the renewable energy system delivers power either to an electric grid, or to a grid-connected building or facility to meet electric load. It does not model thermal energy systems that meet a thermal process load. As mentioned in \cite{NRELBlair}, SAM does not model isolated or off-grid power systems, and does not model systems with electricity storage batteries.
%
%\subsection{HOMER}
%As defined in \cite{HOMER}, actually is a set of two tools: HOMER Legacy and HOMER Pro. HOMER is an acronym for Hybrid Optimization Model for Multiple Energy Resources. HOMER Legacy is the original HOMER software version created at the National Renewable Energy Laboratory (NREL). HOMER Legacy is a free computer model that simplifies the task of evaluating design options for both off-grid and grid-connected power systems for remote, stand-alone, and distributed generation applications. HOMER's optimization and sensitivity analysis algorithms allow the user to evaluate the economic and technical feasibility of a large number of technology options. Since 2016 HOMER Legacy can be found at HOMER web site, but only available for students and nonprofit organizations, as defined in \cite{HOMER}, and has not support available. At the short-time only the commercial version will remain.
%
%The commercial version (paid), called HOMER Pro, as defined in \cite{Swarnkar}, is a tool for optimizing micro-grid design in all sectors, from village power and island utilities to grid-connected campuses and military bases. HOMER Pro put together three tools in one product: optimization, simulation, and sensitivity analysis. HOMER Pro provides the detailed rigor of chronological simulation and optimization in a model that is intended to be easy to use. It is adaptable to a wide variety of projects. For a village or community-scale power system, HOMER can model both the technical and economic factors involved in the project. For larger systems, HOMER can provide an overview that compares the cost and feasibility of different configurations. Chronological simulation is essential for modeling variable resources, such as solar and wind power and for combined heat and power applications, where the thermal load is variable. HOMER's sensitivity analysis helps determine the potential impact of uncertain factors such as fuel prices or wind speed on a given system. 
%
%\subsection{RETScreen}
%As mentioned in \cite{Pradhan}, RETScreen is a decision-support tool designed to help decision makers and energy professionals to evaluate the financial viability of renewable energy, energy efficiency, and/or co-generation projects.
%
%RETScreen models various types of renewable energy technologies (RETs), allowing for comparisons between technology options. The software can be used to evaluate benefits from both clean energy production from power generation projects and savings through energy efficiency projects, accounting for project costs, greenhouse gas emission reductions, and financial risk. The software is freely distributed (but with restrictions to save work or print), and had three different versions:
%
%\begin{itemize}
%\item RETScreen 4 (discontinued, requires Microsoft Excel to run); 
%\item RETScreen Software Suit, which includes the RETScreen 4 and a Windows-based graphical software that allows project owners to verify the ongoing energy performance of their facilities (discontinued in 2013);
%\item And the actual (2016) RETScreen Expert, which allows users to evaluate energy investments over an entire project life-cycle (including benchmarking, feasibility, and performance analysis) in a fully integrated way, and within one software platform. However, this version works is only Windows-based. This version has a complete paid version, in an annual subscription way.
%\end{itemize}
%
%As described by \cite{Pradhan}, RETScreen performs a five-step standard analysis: setting and site conditions; energy model; cost analysis; emission analysis, financial analysis, sensitivity, and risk analysis. It is developed and maintained by the Government of Canada, through the CanmetENERGY Varennes Research Centre of Natural Resources; in collaboration with: NASA; Renewable Energy and Energy Efficiency Partnership (REEEP); United Nations Environment Programme (UNEP), and the Global Environment Facility (GEF). RETScreen is available in 36 languages; it is a multi-awarded tool, and includes equipment databases for components manufactured and available worldwide.
%
%\subsection{Hybrid2}
%The Hybrid2 software package, as described in \cite{Mills}, is a user-friendly tool to perform detailed long-term performance and economic analysis on a wide variety of hybrid power systems. Hybrid2 is a probabilistic/time series computer model, using time series data for loads, wind speed, solar insolation, and temperature; and the power system is designed or selected by the user, in order to predict the performance of the hybrid power system. Variations in wind speed and in load within each time step are factored into the performance predictions. The code does not consider short-term system fluctuations caused by system dynamics or component transients. This program is not supported anymore and according to \cite{UMASS}, probably after the user performs the free download of the tool, it will not work on Windows platforms later than Windows XP, what is a limitation.
%
%Table~\ref{table:softwares} summarizes the off-the-shelf tools employed here, where 
As highlights, only Hybrid2 does not have technical support; HOMER and Hybrid2 perform off-grid system with battery backup analysis. %all the tools perform solar photovoltaic analysis; 
Additionally, HOMER and RETScreen include economical analysis or even optimization-sensitive analysis. However, commercial version of those tools, called RETScreen Expert and HOMER Pro, are available only for Microsoft Windows and the annual subscription typically range from US\$504.00 to US\$657.00.

\begin{table}[!t]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Comparative coverage of reference software}
\label{table:softwares}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{c | c | c | c | c | c}
\hline
\hline
Characteristic  & \rotatebox{90}{PVWatts} & \rotatebox{90}{SAM} & \rotatebox{90}{HOMER} & \rotatebox{90}{RETScreen} & \rotatebox{90}{Hybrid2}\\
\hline
\hline
Support & X & X & X & X &  \\
\hline
Off-grid systems &   &   & X & X & X\\
\hline
Hybrid systems &  &  & X & X & X\\
\hline
Photovoltaics & X & X & X & X & X\\
\hline
Batteries &  &  & X &  & X\\
\hline
\makecell{Main technical (T) \\ or economical(E)} & T & T & E & E & T \\
\hline
Optimization &  &  & X & X &  \\
\hline
Sensitive analysis &  &  & X & X & \\
\hline
\hline
\end{tabular}
\end{table}

%-------------------------------------------------------------------
%\subsection{Paper Proposal x Reference Tools}
%-------------------------------------------------------------------
%
In this study, only HOMER has the minimum features for a comparative with our proposed verification approach. That evaluation is presented in Section~\ref{sec:results_indeed}.%All tools need some parameters inherently from manufacturer's catalog.
%, so the project starts with manufacturers and integrators tool to define the basic items of the project: panels, inverters, controllers, and batteries. 
%After that, the potential solution is analyzed in another tool to simulate or even optimize the proposed solution. Therefore, that is 
%Thus, the main challenge here is to demonstrate the application of software model checking to formally verify a stand-alone PV solution, thus proving that this approach is more effective and complete than other state-of-the-art tool such as HOMER Pro. The comparative evaluation between HOMER Pro and our approach is presented in Section~\ref{sec:results_indeed}.
%
%-------------------------------------------------------------------
\subsection{Component models for a stand-alone PV system }
%-------------------------------------------------------------------
%The main purpose of this section is to describe the models for the elements of a stand-alone PV system: PV generator, battery, controller, inverter, and load. 

>>>>

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{List of Abbreviations, Units and NomenclatureList of Abbreviations, Units and Nomenclature}
\end{center}

\noindent $a$ Ideality or quality factor \\
AC Alternating current \\
$C_{10}$ 10h of rated battery capacity \\
$C_{bank}$ Total capacity of the battery bank \\
CBMC C Bounded Model Checker \\
$ cos \varphi $ Power factor of the inverter \\
CPA Configurable Program Analysis \\
CPAchecker Configurable Program Analysis Checker \\
DC Direct current \\
$DOD$ Depth of discharge ($DOC=1-SOC$) \\
$E_{consumption}$ Energy consumption \\
$E_{corrected}$ Energy corrected \\
$E_{p}$ Energy produced by one panel \\
ESBMC Efficient SMT-based Bounded Model Checker \\
$G$ Solar irradiance \\
$ G_{ref} $ Solar irradiance at STC ($1000 W/m^{2}$) \\
$I$ Current \\
$I_{0}$ Reverse saturation current \\
$ I_{AC} $ Current that the inverter supply to the house \\
$I_{c,min}$ Charge controller minimum current \\
$ I_{DC} $ Current required by the inverter from the DC source \\
$ I_{in} $ Input current \\
$I_{load}$ Current that feed the load \\
$ I_{out} $ Output current \\
$I_{ph}$ Photocurrent \\
$ I_{ph,ref} $ Photocurrent at STC \\
$I_{pv}$ Current from the PV system \\
$ I_{sc} $ Short-circuit current \\
$ I_{sc,amb} $ Ambient short-circuit current \\
$ I_{sc,ref} $ Short-circuit current at STC \\
$ k_{B} $ Boltzmann constant ($ 1.3806503\times10^{-23}J/K $); \\
$MAX_{AC,ref}$ Peak power that the inverter can support \\
$N$ Number of series-connected cells \\
$N_{B}total$ Sized number of batteries \\
$N_{BP}$ Sized number of parallel batteries \\
$N_{BP}min$ Total minimum number of parallel batteries \\
$N_{BS}$ Sized number of series batteries \\
$N_{BS}min$ Total minimum number of series batteries \\
NOCT Nominal Operating Cell Temperature \\
$N_{PS}$ Sized number of panels in series \\
$N_{PSmin}$ Total number of panels in series \\
$N_{PP}$ Sized number of panels in parallel \\
$N_{PPmin}$ Total umber of panels in parallel \\
$ N_{TP} $ Sized number of panels \\
$N_{TPmin}$ Total minimum number of needed solar panels \\
PV Photovoltaic \\
P Power \\
$P_{in}$ Input power \\
$P_{out}$ Output power \\
$P_{surge}$ Maximum peak power demanded by house for few seconds \\
$q$ Absolute value of the electron's charge ($ -1.60217646\times10^{-19}C $) \\
SAT Satisfiability or Boolean satisfiability problem \\
SAT When returns a failure from the verification tool \\
SMT Satisfiability modulo theories \\
$ SOC $ Battery state of charge \\
STC Standard Test Condition (reference state) \\
$T$ Temperature \\
$ T_{air} $ Ambient temperature \\
$ T_{ref}$ Temperature at STC ($=298.15 K$ or $25^{o}C$) \\
UNSAT Condition when verification tool succeeds (could not reach an error) \\
V Voltage \\
$ V_{AC}  $ Voltage that the inverter supply to the house \\
$V_{c}$ Battery charge voltage  \\
$V_{d}$ Battery discharge voltage \\
$ V_{DC} $ Input voltage to the inverter delivered by the DC source \\
$ V_{in} $ Input voltage \\
$ V_{oc} $ Open-circuit voltage \\
$ V_{oc,ref} $ Reference open-circuit voltage at STC \\
$ V_{out} $ Output voltage \\
$ V_{system} $ DC voltage of the bus \\
$V_{T}$ Thermal voltage ($ V_{T}=k_{B}T/q $) \\
$\Delta T$ Temperature variation ($\Delta T=T-T_{ref} $) \\
$ \mu_{I} $ Short-circuit current temperature coefficient ($A/K$) \\
$ \mu_{V} $ Open-circuit voltage temperature coefficient ($ V/K $) \\
$\eta_{b}$ Battery efficiency \\
$\eta_{c}$ Charge controller efficiency \\
$\eta_{i}$ Inverter efficiency \\
$\eta_{p}$ PV panel efficiency \\
